
# Exploration analysis

In order to have a global overview about our table, we are going to have a zoom on each differents table.

## Police Employee Demographics

We start to analyse the dataset of the police officers. We would like to know the proportion of the ethnicities among police officers.We observe in the figure \@ref(fig:distribution-employee) that male gender and the white race are in majority. This distribution is captured in the current year of 2016.

```{r distribution-employee, fig.cap="Distribution of the employee", fig.align='center',warning=FALSE, message=FALSE, echo = FALSE}


Employee_demographics %>% 
  count(Gender,Race) %>%
  ggplot(aes(x=Race,y=n, fill= Race)) +
  geom_bar(stat = "identity") +
  facet_grid(~Gender)+
  theme(legend.position ="bottom") +
  theme(legend.key.size = unit(0.5, "cm"))  +
  labs(
    title = "Majority of male white employee",
    subtitle = "Current of the year 2016",
    x= "Race",
    y="Number of employee"
    ) 



```

Now our interest is to know which positions are the most occupied and by whom. 
Looking at the distribution of the job class in the figure \@ref(fig:distribution-employee), we note that almost all the police positions are attributed to white male people. It is an interesting point for a further analysis.

```{r distribution-job-employee, fig.cap="Distribution of the job", fig.align='center',warning=FALSE, message=FALSE, echo = FALSE}

Employee_demographics %>%
  group_by(`Job Class Description`) %>%
  count(`Job Class Description`, Gender, Race) %>% 
  arrange(desc(n)) %>%
  rename("Number of employee"=n)%>%
  head(7) %>%
  kable(
    caption= "Profession of the police officers by frequency", digits=2) %>%
  kable_styling(
    bootstrap_options = "striped"
    ) 


```


We found interesting to focus on the above job classes, which work in the field. We filtered thus Job Classes Despcription to obtain a subset including the post that figure in the table \@ref(fig:distribution-employee) as the more reccurent.

According to some studies run in the field, a chronic stress caused by police work is an explanation of the use of extralegal force [see @Griffin2003, pp.3--21]. In order to set this assumption, we run two investigations: Figure  \@ref(fig:distribution-serviceyear) depicts the modes of the service lengths among genders for the job cited right above. 

We observe that the service length is slightly longer for men than for women. The median ongoing career length for men is 15 years, whereas the women's is approximatively 6 years shorter (-40%). Even if newcomers lower the ongoing service length, such as the outlier in men's category, we observe that men are thus exposed on a longer period to a chronic stress, which is inherent to the job of police officer [More into details @Purba2019]

```{r distribution-serviceyear, fig.cap="Distribution of the police job", fig.align='center',warning=FALSE, message=FALSE, echo = FALSE}


Employee_demographics %>%
   filter( `Job Class Description` == c("POLICE OFFICER I","POLICE OFFICER II", "POLICE OFFICER III","POLICE OFFICER IV", "POLICE SERGEANT", "POLICE LIEUTENANT", "RESERVE POLICE OFFICER")) %>%
  group_by(Gender) %>%
  ggplot(aes(Gender,`Service Years`, color= Gender)) +
  geom_boxplot() +
  labs(
    title = "Women's service is shorter than the men: men's career are almost quite the double",
    subtitle = "Glimpse of the service length in the Chapel Hill Police District among gender",
    x = "Police officers' gender",
    y = "Service length",
    col = "Gender")


```


Considering now the gender among all categories of job classes, we observe that the police officers are mostly male teams, \@ref(fig:distribution-serviceyear). Out of the four grades, there are only 4 women. Women occupy mostly administrative jobs, not linked to the process nor decision making about arrestation. In this way, we can keep in mind that our analysis would focus mainly on male officers.


```{r distribution-serviceyear-employee, fig.cap="Distribution of the job by gender", fig.align='center',warning=FALSE, message=FALSE, echo = FALSE}
#Comupting the mean for the next plot
#Comupting the mean for the next plot
meanM<- Employee_demographics%>%
  group_by(Gender) %>%
  summarize(mean_service = mean(`Service Years`))

Employee_demographics %>%
  ggplot(aes(`Job Class Description`,`Service Years`)) + 
  geom_point(aes(color = Gender)) +
  geom_hline(yintercept = 14.24603, color = "darkblue") +
  geom_hline(yintercept = 12.02083, color = "pink") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size =6)) +
   labs(
     title = "Female employees are more likely to be in administrative positions",
     subtitle = "The average service is slighty longer for men",
     y="Year of service"
     )

```

In order to contextualize this information, we had to know when those police officers started their career. We computed easily the age of enrollment for all positions (not only the job classes we filtered before), which is the result of their age minus their service length. The age at which the officers began their career is an important information, because young people are more mentally vulnerable than older ones [more details @Thomas2016). Starting one's career at a young age and being exposed to a chronic stress and traumatic events have more impact on the subject due to one's age. 

Among filtered job categories, the starting age is under 30 years old in average, as we observe in the \@ref(fig:distribution-serviceyear). All officers were thus early exposed to chronic stress, and more longer if they gained grades. The hierarchy cumulates more years of experience, what is intuitive because of the system of reward and the legitimization to obtain this position: the contrary would have been surprising.The hierarchy is not necessarily playing a direct role in the deicion making of arrestations. Yet, they give orders and guidelines to the police officers, installing a certain culture and mindset to the troops. 
Police officers' starting age takes part in the same interval than for their superior : between 25 and 30 years old. We have two police officers above this age, what could be explain by either a late career or a change in district. However, it seems that to reach the fourth level in police officer scale, the experience is a necessary element. Those policemens do not have less than 13 years old of experience, which is a large amount of time spent, by assumption, in streets and on missions.
Reserve police officers are experienced policemen kept as an option for mission.

For now, we have to ascertain if only those job classes follow the trend of young enrollment or if it is a general tendancy respected by all the human ressources of the CHPD.

```{r distribution-serviceyear-2, fig.cap="Distribution of the job by gender", fig.align='center',warning=FALSE, message=FALSE, echo = FALSE}


Employee_demographics %>%
  filter( `Job Class Description` == c("POLICE OFFICER I", "POLICE OFFICER II",  "POLICE OFFICER III","POLICE       OFFICER IV", "POLICE SERGEANT", "POLICE LIEUTENANT", "RESERVE POLICE OFFICER")) %>%
  mutate(AGEBEGINNING = Age - `Service Years`) %>%
  ggplot(aes( x = `Job Class Description`, y = AGEBEGINNING,color = `Job Class Description`))+
  geom_boxplot() +
  labs(
    title = "Beyond the early age of enrollment among classes, we note that Police officer's \ngrade is based on service length",
    subtitle = "Age of enrollment and service length per job class in CHPD",
    x = "Job Class",
    y = "Age of enrollment",
    caption = "data from Employee_demographics.csv")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```


To match the previous information with race, we observe in this table \@ref(fig:distribution-serviceyear-race) that white employees are generally more experienced due to a longer career.

```{r distribution-serviceyear-race, fig.cap="Distribution of the job by race", fig.align='center',warning=FALSE, message=FALSE, echo = FALSE}


Employee_demographics %>%
  ggplot(aes(`Job Class Description`,`Service Years`)) + 
  geom_point(aes(color = Race)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size =6)) +
   labs(
     title = "White race has generally a higher service year",
     y="Year of service"
     )
 
```

## Victims' socio-demographics 
Now, we are going to go through the victims data. We wanted to know the annual number of victims and their characteristics.  As a reminder, the data displayd observations from 2010 to 2019.
In Figure \@ref(fig:Victim-distribution), there is a clear distinction between black and white race: white victims outnumber easily black victims.The gender seems not to be significantly important since both graphs look like each other. 
We keep in mind that gender is not a help, but race plays a great role to profile the victim, among other variables.

```{r Victim-distribution, fig.cap="Distribution about the type of the victim", fig.align='center',warning=FALSE, message=FALSE, echo = FALSE}
Incident_reports= Incident_reports %>%
                  mutate(month=month(`Date of Occurrence`), year= year(`Date of Occurrence`))

Incident_reports %>%
  filter(year>=2010) %>%
  count(`Victim Race`, `Victim Gender` , year) %>%
  filter(`Victim Gender`!="NA") %>%
  ggplot(aes(x=year,y=n, fill= factor(`Victim Race`), col= `Victim Race`)) +
  geom_line() +
  facet_grid(~`Victim Gender`)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(legend.position ="bottom") +
  theme(legend.key.size = unit(0.5, "cm"))  +
  labs(title = "White people are the major victim ",
       subtitle = "No significant differenent between the gender",
       x= "Year",
       y="Number of Victim") +
  scale_x_continuous( breaks = seq(2010,2019,by=1))
```



Now that we know the distribution of victims, it would be interesting to have a glimpse over the type of issue they suffered from. 
Therefore, the table \@ref(tab:Table-victim-crime) demonstrates the most frequent complaints involve theft. The third and fourth most frequent issues reported are "misc officier in" and "domestic distur".


```{r Table-victim-crime),warning=FALSE, message=FALSE, echo = FALSE}

Incident_reports %>%
  group_by(`Reported As`) %>%
  count(`Reported As`,`Victim Race`,`Victim Gender`) %>%
  arrange(desc(n)) %>%
  rename("Number of victim"=n) %>%
  na.omit() %>%
  head(10) %>%
   kable(
    caption= "Purpose of the crime", digits=2) %>%
  kable_styling(
    bootstrap_options = "striped"
    ) 


```




In the figure \@ref(fig:delay-occurence-report), we would like to know if there is a difference in process time for registering the complaint and the the arrestation process among race of victims. Even if there are some variations depending on the complexity of the case and the inherent time of investigation depending on the type of case, variation we cannot control, it would be interesting if there are anyway a global difference since our panel of 82'525 observations would somehow pool this variance.

Outliers are the typical cases showing this uncontrollable variation, we would neglect them and filter our dataset to observe only the delays inferior to 2000 days (more than 3 years). We observe that the time dedicated to investigation and arrestation for cases of white people is slightly higher than this time for black people. We would discard the assumption of a negligence in the treatment of cases related to black people.


```{r delay-occurence-report, fig.cap="Bloxplot about the delay between the occurence and the report", fig.align='center',warning=FALSE, message=FALSE, echo = FALSE}
Incident_reports%>%
  group_by(`Victim Race`)  %>%
  mutate(DELAYReported = `Date of Report`- `Date of Occurrence`, DELAYArrested = `Date Found`- `Date of Occurrence`) %>%
  filter( DELAYArrested < 20000 & DELAYArrested > 100) %>%
  ggplot( aes(x= `Victim Race`,y= DELAYArrested, color=`Victim Race`)) +
  geom_boxplot() +
  labs(
    title = "Higher delay between the occurence and the date of the report for the black race",
    subtitle = "More outliers for the white race",
    y="Delay between the occurence and the report" 
    )


```

## Arrestees' socio-demographics

In the data "Police arrested made", there are all the data of the criminals for their arrest from 2010 to 2019. The first thing we wished to observe is the number of arrests per month and per year (for a better visualization) classified by race of the person arrested.

In Figures \@ref(fig:Police-arrested-made-Black-White-month) and \@ref(fig:Police-arrested-made-Black-White-year), we observe that from 2015, the number of arrestations for black people increases.

The below figures show the number of arrestations by race and we note an equivalent proportion of black and white arestees. However, if we confront this information with the socio-demographic data of White Chapel in 2010, the black population represented only 9.9%. According to the data of [2020 World Population by Country(updated daily)](https://worldpopulationreview.com/) displaying the population of Chapel Hill and its population growth, the population growth between 2010 and 2018 was 6.24%, so let's assume that the population of black people increased by the same percentage, and the variation would be +/- 6.24% [0% or 12.48%] It would lead that in 2018, on year before our data, we should have 10.5% of black people in White Chapel Hill or a percentage included in [9.9% ; 11.17%]. Here, if we had to respect the proportion in the arrestee sample according to the white people number of arestees, it should be around 386 black aresteesin order to be representative of the whole population. The entered 3'651 black arrestees represent 945% of this extrapolation.

Since 2015, the trend is inverted and black arrestees sometimes outnumbers white people. Sometimes, the difference is shocking, for instance in March-April and July-August 2016 or June-July 2018. These differences are peaks that could be related to the meeting and strikes related to black shooting occuring in the year and to repeal  a law (HB2), as the articles enclosed in the appendix relate.[@Newobserver, title "Hundreds block downtown Chapel Hill to protest state’s HB2 law"]

We do neglect the possibility that mass movements could as well explain unusual peaks. However, as we mentioned earlier, the inverted trend is significantly verified in 2017, where white arrestees were a majority only during two first months. This fact could not be explained by mass movements.

```{r Police-arrested-made-Black-White-month, fig.cap="Evolution of the number of arrestation by month",fig.align='center', warning=FALSE, message=FALSE, echo = FALSE}

Police_arrest_made <-Police_arrest_made %>% 
 mutate(month=month(`Date of Arrest`),year=year(`Date of Arrest`))


Police_arrest_made %>% 
count(`Arrestee's Race`, month, year) %>%
ggplot(aes(x=month,y=n)) +
  geom_line(size=1, aes(color=`Arrestee's Race`)) +
  geom_point(aes(color=`Arrestee's Race`)) +
  facet_grid(~year) +
  stat_summary(fun.y = diff, geom = "bar") +
  theme(legend.position ="bottom")  +
  scale_x_continuous( breaks = c(1, 6, 12))+
  labs(
    title = "Monthly number of arrestations",
    x= "Month",
    y="Number of arrestations"
  )

```



```{r Police-arrested-made-Black-White-year, fig.cap="Evolution of the number of arrestation by year",fig.align='center', warning=FALSE, message=FALSE, echo = FALSE}
##Same plot as before, but only considering the year and not the month


Police_arrest_made %>%
count(`Arrestee's Race`, year,) %>%
ggplot(aes(x=year,y=n)) +
  geom_line(size=1, aes(color=`Arrestee's Race`)) +
  geom_point(aes(color=`Arrestee's Race`)) +
  stat_summary(fun.y = diff, geom = "bar") +
  theme(legend.position ="bottom")  +
   scale_x_continuous( breaks = seq(2010,2019,by=1))+
  labs(
    title = "Annual number of arrestations",
    x= "Year",
    y="Number of arrestation"
  )

 
```


After dividing the arrests by class, we wanted to know the gender of the arrestees and we observed, in the figure \@ref(fig:Arrestation-Male-Female-Black-White), that the number of men is in average 3 times the number of arrested women, both races combined. The male arrestees number remains above the 600 for white and for black (except in 2012, around 580) from 2010 to 2015. Until that date, Black arrestees' number belongs to an approximate interval of [580;800], with a high increase between 2012-2015 (approx. +35%). White arrestees belongs to the interval [600:800]. Both races decreases tremendously afterwards, but the black community always remains higher of +50 numbers of arrestees in average, which denotes well the disparity among races.

In 2015, some institutional policy must have been applied to decrease the number of arrestee preventively or punishingly because the number of arrestees has decreased for both white and black citizens. The trend of a decrease in the number of arrestees is observed one year later for women: in 2016. Women arrestees number is quite stable, oscillating in an interval of [100;200] for white women and [50;150]



```{r Arrestation-Male-Female-Black-White, fig.cap="Evolution of the number of arrestation between male and female",fig.align='center', warning=FALSE, message=FALSE, echo = FALSE}

Police_arrest_made %>%
  count(`Arrestee's Race`, `Arrestee's Gender`, year) %>%
  filter(!is.na(`Arrestee's Race`)) %>%
  ggplot(aes(x=year,y=n, fill= factor(`Arrestee's Race`), col= `Arrestee's Race`)) +
  geom_line() +
  facet_grid(~`Arrestee's Gender`)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(legend.position ="bottom") +
  theme(legend.key.size = unit(0.5, "cm"))  +
  labs(title = "Higher number of arrestation for male people",
       subtitle = "Most of the time: white arrested = female, black arrested = Male",
       x= "Year",
       y="Number of arrestation"
       ) +
  scale_x_continuous( breaks = seq(2010,2019,by=1))
```

And what about their ages?
The age and the race would complete the profile of our arrestees. White arrestees are slightly younger (27-28 yo) considering the median than black people (34-35yo), Even the 1rst and 3rd modes are present a difference of about 5 years, making black arestees older. And we note that the mean age of black arrestees in slightly superior to the white arrestees mean age. The mean ages are respectedly 33 years for black people and 25 for white people, as observed in the figure \@ref(fig:Arrestation-age).

```{r Arrestation-age, fig.cap="Age of the arrested people related to the race", fig.align='center',warning=FALSE, message=FALSE, echo = FALSE}

Police_arrest_made%>%
  ggplot(aes(x=`Arrestee's Race`,y= `Arrestee's Age`, color=`Arrestee's Race` ))+
  geom_boxplot() +
  facet_grid(~`Arrestee's Gender`)+
  labs (
    title = "Median age higher for the black arrested",
    subtitle = "Arrested pepople between 2010 and 2019",
    x= "Arrestee's Race",
    y= "Age"
    )
```


We would now consider the type of arrestations \@ref(fig:Arrestation-type-arrestation). Black arrestees have more chance to be taken into custody than white people (chance until 2 times the white people have in 2011). It's hitting to remark that the heavier measures police officers could apply to arrestees is attributed in those proportions to black people and this difference in treatment among races remains quite stable. Each year verifying the schocking gap between white and black arrestees taken into custody.
 
Since 2012, the black proportion of peolple summoned or cited increased to reach 300 arrestees per year, in 2015 and 2016. The curve for white arrestees depicts a yearly decrease, starting in 2010 with a number higher than 400, reaching a flat stage for about 3 years (2012-2015) and the decreased has accelerated since this moment. However, we note that in this category as well, the curves of respectively black and white people arrested have become mixed for the past 5 years, making this sample as well not representative for the entire population of Chapel Hill. 

The same comment could be made for arrestees categorized "on view"", with a tinier gap between white and black people, but always at the cost of black persons.

```{r Arrestation-type-arrestation, fig.cap="Number of arrestation separeted by type and race", fig.align='center',warning=FALSE, message=FALSE, echo = FALSE}

##Analyze the type of arrestation taking into account the race

Police_arrest_made %>%
  count(`Arrestee's Race`, `Type of Arrest`, year)%>%
  filter(!is.na(`Type of Arrest`)) %>%
  ggplot(aes(x=year,y=n, fill= factor(`Arrestee's Race`), col= `Arrestee's Race`)) +
  geom_line() +
  facet_grid(~`Type of Arrest`)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(legend.position ="bottom") +
  theme(legend.key.size = unit(0.5, "cm"))  +
  labs(title = "Black arrestees are mostly taken into custody",
       subtitle = "Arrestations by type of arrest",
       x= "Year",
       y="Number of arrestation") +
  scale_x_continuous( breaks = seq(2010,2019,by=1))


```


Victims and arrestees are now confronted in order to now the distribution of conflicts \@ref(tab:victim-arrested). The interracial conflicts are lower than white vs white and black vs black complaints. The proportion of white victims here, neglecting any observations involving other races (Asian, Hispanic, etc.), is a subset closer to the real population. Black victims represent 34.31% of this sample of victims, whereas the real proportion of black is approx. 10% over 100% of population. In spite the 34.31% is clearly a higher rate than the proportion of black victims we could have expected, we won't investigate the potential causes of this great rate. The point here is really to confront racial and interracial conflicts distribution. 


 
```{r victim-arrested,warning=FALSE, message=FALSE, echo = FALSE}

Table_joint<-left_join(Incident_reports, Police_arrest_made, by = c("Date of Occurrence"="Date of Arrest"))


Table_joint_2 <- Table_joint %>%
              select(year.x,`Arrestee's Race`, `Victim Race`) 




table(Table_joint_2$`Victim Race`,Table_joint_2$`Arrestee's Race`) %>%
 kable( 
  caption = "Vertical = Victim  |  Horizontal = Arrested", digits = 2
  ) %>% 
  kable_styling(
  bootstrap_options = "striped"
  )
  

```


## Drivers and Passengers' socio-demographics

Finally, we decided to analyse the traffic arrestation and again to differentiate black to white people. To do so, we chose an alternative method based on percentage: a ratio determining how many black citizens more were arrested for a certain motive of arrestx, scaled by the number of white arrestees, per year.

The most striking information is the ratio is the highest for police investigation. In other words, for each 10 white citizen arrested for investigation, whatever the gender is, at least 16 black persons already have been arrested. Those 60% of difference between those two races clearly signifies a correlated disproportion. However, narrowing down the specific cause of it is impossible and is not part of the present work. 

The reader has to keep in mind the black inhabitants' proportion in demographics dated 2010 in Chapel Hill (10%) and our extrapolation: (+/- 1.2%) per year, with respect to the whole population. The present arrestees sample is absolutely not representative of our population, even if estimated.

In Figure \@ref(fig:Traffic-Ratio-Black-White), if the bar is higher than 50%, it mean that more black persons are arrested and inversely. We note that for the purpose "Investigation", more black are arrested for the both year and independently of  the gender. The feature "Driving while impaired"" concern mostly black males.


```{r Traffic-Ratio-Black-White, fig.cap="More black are arrested",fig.align='center', warning=FALSE, message=FALSE, echo = FALSE}

 Traffic_stop_2 <- Traffic_stop_2 %>% mutate(Pro_Black_white = Black/White)
 Traffic_stop_2$Pro_Black_white[is.nan(Traffic_stop_2$Pro_Black_white)]<-0
 
 
Traffic_stop_2 %>% group_by(Purpose, Gender) %>%
ggplot(aes(x= reorder(Purpose,Traffic_stop_2$Pro_Black_white) , y= Traffic_stop_2$Pro_Black_white)) +
  geom_bar(stat = "identity") +
  facet_grid(Gender~Year) +
  labs(title = "More black arrested for the investigation purpose" ,subtitle = "for the year 2014-2015", x= "Purpose", y= "Ratio Black/White")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
   scale_y_continuous(labels = scales::percent) +
  theme( strip.background  = element_blank(),
                      panel.grid.major = element_line(colour = "black"),
                      panel.border = element_blank(),
                      axis.ticks = element_blank(),
                      panel.grid.minor.x=element_blank(),
                      panel.grid.major.x=element_blank() )

```


To refine the exploration, we want to know what type of arrest prevails per year \@ref(fig:Traffic-stop).

The strict majority among arrestation categories is speed limit violation, which overpass 4'000 arrestations per year, in both 2014 and 2015, and whose gender proportion approximate a 55%-45% (quite homogeneous). Speed violation is a daily risk for drivers, whether it is voluntary done or not. For this reason, speed violation was intuititive first cause of verbalization. 

The second and third causes of arrest concerned violation of vehicle regulations or equipment regulations. Which is interesting to note here is that the first causes of arrests, whose number of occurences exploded in 2014 and 2015, concerns a behavioral cause. Behavioral causes could imply violent reactions from police officers who are/feel agressed, whereas non-behavioral causes (ex. vehicule equipment violation) lead less to violence or other bias in decision making for  arresting the verbalized person(s). 
Other behavioral causes are fewly reported: (safe or other) movement violations, driving whiled impaired, etc. Offensive or violent behaviors are not significantly represented here other than potentially trhough speed violation which, as we previously said, could be unwanted.


```{r Traffic-stop, fig.cap="The higer purpose is the speef of limit",fig.align='center', warning=FALSE, message=FALSE, echo = FALSE}

Traffic_stop_2<-Traffic_stop_2 %>% filter(Purpose != "Total" & Purpose != "FemaleTotal" & Purpose != "MaleTotal")

ggplot(Traffic_stop_2, aes(x=Purpose, y= Traffic_stop_2$`Total By Ethnicity`, fill= Gender)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Year) +
  labs(title = "The main cause of traffic arrestation is speed limit violation" ,subtitle = "For the year 2014-2015", x= "Purpose", y= "Total stop by race") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```




