# Introduction

## Overview and Motivation

Our motivation is to work on the characteristics explaining crime and (im)partiality of the American judicial system applied. In addition to shootings on public places, cliché rift between black/white people often agitates news and press and stirs up American society. For this project we will use data from Chapel Hill Open Data, which is about the crime in this state

By using the CHPD database, far from finding a universal answer applicable to the United States of America in a whole because our sample does not allow us to do so, we have the possibility of considering the socio-demographic characteristics of offenders and police officers over a defined period and time span in order to check whether the clichés of the cleavage can be confirmed or refuted. The use of this database potentially allows us to identify some potentially significant correlations between factors of delinquency or bias among police officers

Our goal is to find some significant correlation between the variables in order to profile the different offenders and see if these profiles infer with the typical characteristics of police profiles. 
Does the concentration of police officer profiles have a link with the profiles of offenders, the offences committed or the satisfaction of the population with the service they receive from the police?
From sociodemographic of the people arrested do stem some insight on some contributing factors of arrestation among a certain type of population or a precise distribution of delinquency among the socio demographics?


## The data

```{r warning=FALSE, message=FALSE, echo = FALSE}


Traffic_stop<- read_delim("data/drivers-and-passengers-searched-by-sex-race-and-ethnicity.csv",";", escape_double = FALSE, trim_ws = TRUE)

 Traffic_stop_2<-read_delim("data/initial-purpose-of-traffic-stop-by-drivers-sex-race-and-ethnicity.csv", ";", escape_double = FALSE, trim_ws = TRUE)
 
Police_arrest_made <- read_delim("data/police-arrests-made.csv",";", escape_double = FALSE, trim_ws = TRUE)
 
Employee_demographics<- read_delim("data/police-employee-demographics.csv",";", escape_double = FALSE, trim_ws = TRUE)

Incident_reports <-  read_delim("data/police-incident-reports-written.csv",";", escape_double = FALSE, trim_ws = TRUE)

```

As mentioned above ,for this assignement we use then Chapel Hill Crime and Police Data.

These sets of datasets will focus on datasets gathered from Chapel Hill Open Data, which is managed and run by the town of chapel hill. This readme is a compilation of information that represents what each file will be about. Most of the datasets can be understood by reading the summary page from the Chapel Hill Open Data [website](https://www.chapelhillopendata.org/explore/?sort=modified&refine.keyword=Police)

We are going to use four differents tables. We didn't have to do any data cleansing. We just renamed the names of a few observations so that the reader could understand what they were about.


### Police Incidents Reports Written
This dataset contains incident reports filed with the Chapel Hill Police Department. Multiple incidents may have been reported at the same time. The most serious incident is shown for each report filed. Data available since January 1, 2010 and updated quarterly.
Observations: 82’525
Variables: 18

For this project, we are going to use the 6 variables present in the table \@ref(tab:Table-Incidents-Reports). Those variables seems to be the most important for our project.

```{r Table-Incidents-Reports, warning=FALSE, message=FALSE, echo = FALSE}
Incident_reports%>%
  select(`Date of Report`,`Date of Occurrence`,`Reported As`,`Victim Age`,`Victim Race`,`Victim Gender`) %>%
  head(5)%>%
  na.omit(Incident_reports) %>%
    kable(caption= "Table incident reports") %>%
  kable_styling(bootstrap_options = "striped")


``` 



### Drivers and Passengers Searched By Sex Race and Ethnicity
This dataset contains data on drivers and passengers searched during traffic stops, broken down by sex, race, and ethnicity.
Observations: 8 
Variables: 14
In the table \@ref(tab:Table-Traffic-Stop), you have an overview of the variables present in this dataset. Further, for our project we are going only to take into account black and white race, as the other seems to be less significant.

```{r Table-Traffic-Stop, warning=FALSE, message=FALSE, echo = FALSE}
Traffic_stop%>%
  select(-`Total By Race`,-TotalStopped,-`Total By Ethnicity`,-`Percent Searched`) %>%
  head() %>%
    kable(caption= "Table Traffic Stop", digits=2) %>%
  kable_styling(bootstrap_options = "striped")

``` 

### Initial Purpose of Traffic Stop by Drivers Sex Race and Ethnicity
This dataset contains data on the initial purpose of traffic stops, broken down by sex, race, and ethnicity.
Observations: 46
Variables:12
The table \@ref(tab:Table-Traffic-Stop-2) is the quite the same as the table \@ref(tab:Table-Traffic-Stop), but we have the purpose of each crime. 

```{r Table-Traffic-Stop-2, warning=FALSE, message=FALSE, echo = FALSE}
Traffic_stop_2 %>%
  select(Year,Purpose,Gender,White,Black) %>%
  head(5) %>%
    kable(caption= "Table Traffic stop 2", digits=2) %>%
  kable_styling(bootstrap_options = "striped")

``` 

### Police Arrests Made
This dataset contains arrests made by the Chapel Hill Police Department. Multiple individuals may have been arrested under the same arrest number during related incidents or arrests. Data available since January 1, 2010 and updated quarterly.
Observations: 16040
Variables: 12

There are all the arrests from 2010 to 2019. We take into account the 8 variables that you can see in the table  \@ref(tab:Table-Police-Arrest-Made),

```{r Table-Police-Arrest-Made, warning=FALSE, message=FALSE, echo = FALSE}
Police_arrest_made %>%
  select(-`Incident Id`,-`Time of Arrest`,-Address, -`Arrestee's Ethnicity`) %>%
  na.omit(Incident_reports) %>%
  head(5) %>%
    kable(caption= "Report of the criminal", digits=2) %>%
  kable_styling(bootstrap_options = "striped")
``` 


### Police Employee Demographics
 This dataset contains demographics information for employees of the Chapel Hill Police Department. Current as of October 2016.
Observations: 174
Variables: 7

In the table \@ref(tab:Table-Police-Employee-Demographics),

```{r Table-Police-Employee-Demographics, warning=FALSE, message=FALSE, echo = FALSE}
Employee_demographics %>%
  select(-`Officer ED No.`) %>%
  head(5) %>%
    kable(caption= "Employee Demographics", digits=2) %>%
  kable_styling(bootstrap_options = "striped")
``` 


```{r warning=FALSE, message=FALSE, echo = FALSE}
  Police_arrest_made <- Police_arrest_made%>% 
   mutate( `Drugs or Alcohol Present` = fct_collapse(`Drugs or Alcohol Present`,
    "Yes"="Y",
    "No"="N",
    "Unknown"="U"
    )) %>%
   mutate(`Arrestee's Race` = fct_collapse(`Arrestee's Race`,
    "Asian"="A",
    "Black"="B",
    "Hispanic"="H",
    "Native American"="I",
    "Other" ="O",
    "Unknown"="U",
    "White"="W")) %>%
  mutate(`Arrestee's Gender`=fct_collapse(`Arrestee's Gender`,
        "Female"="F",
        "Male"="M")) %>%
  mutate(`Type of Arrest`=fct_collapse(`Type of Arrest`,
         "TAKEN INTO CUSTODY"="TAKEN INTO CUSTODY (WARRANT/LP)")) %>%
  filter(`Arrestee's Race`=="Black" | `Arrestee's Race` == "White")

  
Incident_reports<- Incident_reports%>% 
  mutate(`Victim Race` = fct_collapse(`Victim Race`,
    "Asian"="A",
    "Black"="B",
    "Hispanic"="H",
    "Native American"="I",
    "Other" ="O",
    "Unknown"="U",
    "White"="W")) %>%
  mutate(`Victim Gender`=fct_collapse(`Victim Gender`,
    "Female"="F",
    "Male"="M")) %>%
  filter(`Victim Age`<100)  %>%
  filter(`Victim Race`=="Black" | `Victim Race`=="White") 

Employee_demographics<- Employee_demographics %>%
  mutate(Race = fct_collapse(Race,
    "Asian"="A",
    "Black"="B",
    "Hispanic"="H",
    "Native American"="I",
    "Other" ="O",
    "Unknown"="U",
    "White"="W")) %>%
  mutate(Gender=fct_collapse(Gender,
    "Female"="F",
    "Male"="M"))%>%
  filter(Race=="Black" | Race=="White")

Traffic_stop_2 <- Traffic_stop_2 %>%
  mutate(Gender= fct_collapse(Gender,
        "Female"="F",
        "Male" = "M"))

```
