# Introduction

## Overview and Motivation

Our motivation is to work on the characteristics of arrestations explaining a racial bias of the American judicial system applied. In addition to shootings on public places, cliché rift between black/white people often agitates news and press and stirs up American society. For this project we will use data from Chapel Hill Open Data, which compiles observations on police officers, victims and arrestees.

By using the CHPD database, far from finding an universal answer applicable to the United States of America in a whole, because our sample does not allow us to do so, we have the possibility of considering the socio-demographic characteristics of offenders and police officers over a defined  time span in order to check whether the cliché racial targetting can be confirmed or refuted. The use of this database potentially allows us to identify some potentially significant correlations between factors of delinquency or bias among police officers

Our goal is to find some significant correlation between the variables in order to profile the different offenders and see if these profiles infer with the typical characteristics of police profiles. 
From sociodemographic of the people arrested do some insight stem on some contributing factors of arrestation among a certain type of population or a precise distribution of offenders, as well as a homogeneous population of victims or not?


## The data

```{r warning=FALSE, message=FALSE, echo = FALSE}


Traffic_stop_2<-read_delim("data/initial-purpose-of-traffic-stop-by-drivers-sex-race-and-ethnicity.csv", ";", escape_double = FALSE, trim_ws = TRUE)
 
Police_arrest_made <- read_delim("data/police-arrests-made.csv",";", escape_double = FALSE, trim_ws = TRUE)
 
Employee_demographics<- read_delim("data/police-employee-demographics.csv",";", escape_double = FALSE, trim_ws = TRUE)

Incident_reports <-  read_delim("data/police-incident-reports-written.csv",";", escape_double = FALSE, trim_ws = TRUE)

```


```{r warning=FALSE, message=FALSE, echo = FALSE}
  Police_arrest_made <- Police_arrest_made%>% 
   mutate( `Drugs or Alcohol Present` = fct_collapse(`Drugs or Alcohol Present`,
    "Yes"="Y",
    "No"="N",
    "Unknown"="U"
    )) %>%
   mutate(`Arrestee's Race` = fct_collapse(`Arrestee's Race`,
    "Asian"="A",
    "Black"="B",
    "Hispanic"="H",
    "Native American"="I",
    "Other" ="O",
    "Unknown"="U",
    "White"="W")) %>%
  mutate(`Arrestee's Gender`=fct_collapse(`Arrestee's Gender`,
        "Female"="F",
        "Male"="M")) %>%
  mutate(`Type of Arrest`=fct_collapse(`Type of Arrest`,
         "TAKEN INTO CUSTODY"="TAKEN INTO CUSTODY (WARRANT/LP)")) %>%
  filter(`Arrestee's Race`=="Black" | `Arrestee's Race` == "White")

  
Incident_reports<- Incident_reports%>% 
  mutate(`Victim Race` = fct_collapse(`Victim Race`,
    "Asian"="A",
    "Black"="B",
    "Hispanic"="H",
    "Native American"="I",
    "Other" ="O",
    "Unknown"="U",
    "White"="W")) %>%
  mutate(`Victim Gender`=fct_collapse(`Victim Gender`,
    "Female"="F",
    "Male"="M")) %>%
  filter(`Victim Age`<100)  %>%
  filter(`Victim Race`=="Black" | `Victim Race`=="White") 

Employee_demographics<- Employee_demographics %>%
  mutate(Race = fct_collapse(Race,
    "Asian"="A",
    "Black"="B",
    "Hispanic"="H",
    "Native American"="I",
    "Other" ="O",
    "Unknown"="U",
    "White"="W")) %>%
  mutate(Gender=fct_collapse(Gender,
    "Female"="F",
    "Male"="M"))%>%
  filter(Race=="Black" | Race=="White")

Traffic_stop_2 <- Traffic_stop_2 %>%
  mutate(Gender= fct_collapse(Gender,
        "Female"="F",
        "Male" = "M"))

```

As mentioned above, for this assignement we use then Chapel Hill Crime and Police Data.

These sets of datasets will focus on gathered information by Chapel Hill on the town and ts citizen. This readme is a compilation of information that represents what each file will be about. Most of the datasets can be understood by reading the summary page from the Chapel Hill Open Data: [website](https://www.chapelhillopendata.org/explore/?sort=modified&refine.keyword=Police)

We are going to use four differents tables. We didn't have to do any data cleansing. We only renamed the levels of a few observations for the understanding of the reader (ex: "B" renamed as "Black").


### Police Employee Demographics
This dataset contains demographics information for employees of the Chapel Hill Police Department. Current as of October 2016.
Observations: 174
Variables: 7

In the table \@ref(tab:Table-Police-Employee-Demographics), we are going to use the 5 followings variables.

```{r Table-Police-Employee-Demographics, warning=FALSE, message=FALSE, echo = FALSE}
Employee_demographics %>%
  select(-`Officer ED No.`,-Status) %>%
  head(5) %>%
    kable(caption= "Employee Demographics", digits=2) %>%
  kable_styling(bootstrap_options = "striped")
``` 


### Victims' socio-demographics

This dataset, named "Police Incidents Reports", contains incident reports filed by the CHPD. Multiple incidents may have been reported at the same time. The most serious incident is shown for each report filed. Data available since January 1, 2010 and updated quarterly.
Observations: 82’525
Variables: 18

For this project, we are going to use the 6 variables appearing in the table \@ref(tab:Table-Incidents-Reports). Those variables seems to be the most important for our project.

```{r Table-Incidents-Reports, warning=FALSE, message=FALSE, echo = FALSE}
Incident_reports%>%
  select(`Date of Report`,`Date of Occurrence`,`Reported As`,`Victim Age`,`Victim Race`,`Victim Gender`) %>%
  head(5)%>%
  na.omit(Incident_reports) %>%
    kable(caption= "Table victims' socio-demographics") %>%
  kable_styling(bootstrap_options = "striped")


``` 

### Arrestees' socio-demographics
This dataset contains arrests made by the Chapel Hill Police Department. Multiple individuals may have been arrested under the same arrest number during related incidents or arrests. Data available since January 1, 2010 and updated quarterly.
Observations: 16040
Variables: 12

There are all the arrests from 2010 to 2019. We take into account the 7 variables that you can see in the table  \@ref(tab:Table-Police-Arrest-Made),

```{r Table-Police-Arrest-Made, warning=FALSE, message=FALSE, echo = FALSE}

Police_arrest_made %>%
  select(-`Incident Id`,-`Time of Arrest`,-Address, -`Arrestee's Ethnicity`, -`Primary Charge`)%>%
  na.omit(Incident_reports) %>%
  head(5) %>%
    kable(caption= "Arrestees' socio-demographics", digits=2) %>%
  kable_styling(bootstrap_options = "striped")
``` 


### Drivers and Passengers' socio-demographics
This dataset contains data on the initial purpose of traffic stops, entered in the data base by gender, race, and ethnicity.
Observations: 46
Variables:12

In this dataset, we are going to use the 6 followings variables.
```{r Table-Traffic-Stop-2, warning=FALSE, message=FALSE, echo = FALSE}
Traffic_stop_2 %>%
  select(Year,Purpose,Gender,White,Black) %>%
  head(5) %>%
    kable(caption= "Drivers and Passengers' socio-demographics", digits=2) %>%
  kable_styling(bootstrap_options = "striped")

``` 




